Report: Illustration Results

### 2023-02-24 IllustrationResultsRenderReport.Rmd
:sparkles: cluster and performance results expl added

```{r}
#| label: render cluster evaluation
#| include: true
#| results: 'asis'

rmarkdown::render(
  input = "sections/06ClusterEvaluation.Rmd",
  output_dir = outputDir
)
```

>>>>

## CLuster Performance

```{r}
#| label: render cluster performance
#| include: true
#| results: 'asis'

rmarkdown::render(
  input = "sections/06ClusterEvaluationPerformance.Rmd",
  output_dir = outputDir
)
```

## CLuster Interpretation

```{r}
#| label: render cluster interpretation
#| include: true
#| results: 'asis'

rmarkdown::render(
  input = "sections/07ClusterEvaluationInterpretation.Rmd",
  output_dir = outputDir
)
```

### 2023-02-24 latex-results/05FeatureClustering.tex
:sparkles: cluster and performance results expl added

It is this \(WCSS\) that becomes the objective function to be minimized,
by iteratively moving features from one cluster to another
\citep{hartigan1979}. In particular, the algorithm (1) calculates the
cluster centroids of the initial partitioning, (2) checks whether any
feature has a centroid that is closer than that of the currently
assigned cluster (3) updates the centroids based on any reassigned
features, and then iterates between steps two and three until \(WCSS\)
is minimized (i.e., locally optimal convergence) or a maximum number of
iterations is reached. Given the iterative nature of the algorithm, the
initial partitioning is often important because the algorithm might
arrive at a suboptimal clustering where the \(WCSS\) cannot be further
reduced by moving any feature to another cluster, despite a better
solution existing (i.e., a local minimum). It is, therefore often
recommended to run the k-means clustering with several different
starting positions.

In our case we used the Elbow and the Silhouette method to determine the
recommended number of \(k\) clusters, which for both methods was two
clusters (see Supplemental Material A). We then entered the
participants' PC-scores from the feature reduction step into the k-means
algorithm. To avoid local minima we used 100 random initial centroid
positions. In the final solution the k-means analysis assigned 80
participants to cluster 1 and 76 participants to cluster 2.

>>>>

It is this \(WCSS\) that becomes the objective function to be minimized,
by iteratively moving features from one cluster to another
\citep{hartigan1979}. In particular, the algorithm (1) calculates the
cluster centroids of the initial partitioning, (2) checks whether any
feature has a centroid that is closer than that of the currently
assigned cluster (3) updates the centroids based on any reassigned
features, and then iterates between steps two and three until \(WCSS\)
is minimized (i.e., locally optimal convergence) or a maximum number of
iterations is reached \citep{jain2010}. Given the iterative nature of
the algorithm, the initial partitioning is often important because the
algorithm might arrive at a suboptimal clustering where the \(WCSS\)
cannot be further reduced by moving any feature to another cluster,
despite a better solution existing
\citep[i.e., a local minimum;][]{timmerman2013}. It is, therefore often
recommended to run the k-means clustering with several different
starting positions.

In our case we entered the participants' PC-scores from the feature
reduction step into the k-means algorithm. Because we did know the
underlying number of clusters within our sample, we calculated the
cluster solutions for \(k =\) 2 through 10. To avoid local minima we
used 100 random initial centroid positions for each run. Each of the 9
cluster solutions converged within the iteration limit. In the next step
we will then evaluate which of the extracted cluster solutions offers
the best fit with the data.

### 2023-02-24 sections/05FeatureClustering.Rmd
:sparkles: cluster and performance results expl added

# prep data
scaled_pc <- pca.out #%>% scale

# Elbow method
fviz_nbclust(scaled_pc, kmeans, method = "wss") +
  geom_vline(xintercept = 2, linetype = 2)+
  labs(subtitle = "Elbow method") +
  theme_Publication()
# Silhouette method
fviz_nbclust(scaled_pc, kmeans, method = "silhouette")+
  labs(subtitle = "Silhouette method") +
  theme_Publication()
```

@@

kmeansOut <- kmeans(scaled_pc, centers = 2, nstart = 100)
```

```{r}
#| label: Plot Convex Hulls

fviz_cluster(kmeansOut, geom = "point", data = scaled_pc, ggtheme = theme_Publication())
fviz_cluster(kmeansOut, geom = "point", data = z_data, ggtheme = theme_Publication())
```

<!-- \Question{\textcolor{cyan}{It is good practice to standardize the input of the k-means algorithm. However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...). If I actually standardize the PC-score, we get really weird k-means results. The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}  -->

<!-- I had originally written:  -->
<!-- "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores."  -->
<!-- but I went back to the 'raw' PC-scores now. But we should discuss this.}} -->

@@

It is this $WCSS$ that becomes the objective function to be minimized, by iteratively moving features from one cluster to another \citep{hartigan1979}. In particular, the algorithm (1) calculates the cluster centroids of the initial partitioning, (2) checks whether any feature has a centroid that is closer than that of the currently assigned cluster (3) updates the centroids based on any reassigned features, and then iterates between steps two and three until $WCSS$ is minimized (i.e., locally optimal convergence) or a maximum number of iterations is reached. Given the iterative nature of the algorithm, the initial partitioning is often important because the algorithm might arrive at a suboptimal clustering where the $WCSS$ cannot be further reduced by moving any feature to another cluster, despite a better solution existing (i.e., a local minimum). It is, therefore often recommended to run the k-means clustering with several different starting positions.

In our case we used the Elbow and the Silhouette method to determine the recommended number of $k$ clusters, which for both methods was two clusters (see Supplemental Material A). We then entered the participants' PC-scores from the feature reduction step into the k-means algorithm. To avoid local minima we used 100 random initial centroid positions. In the final solution the k-means analysis assigned `r table(kmeansOut$cluster)[1]` participants to cluster 1 and `r table(kmeansOut$cluster)[2]` participants to cluster 2.

>>>>

# prep data
scaled_pc <- pca.out #%>% scale
```

@@

kmeansRes <- list()
for (i in 2:10) {
  kmeansRes[[i-1]] <- kmeans(scaled_pc, centers = i, nstart = 100)
}
```

@@

It is this $WCSS$ that becomes the objective function to be minimized, by iteratively moving features from one cluster to another \citep{hartigan1979}. In particular, the algorithm (1) calculates the cluster centroids of the initial partitioning, (2) checks whether any feature has a centroid that is closer than that of the currently assigned cluster (3) updates the centroids based on any reassigned features, and then iterates between steps two and three until $WCSS$ is minimized (i.e., locally optimal convergence) or a maximum number of iterations is reached \citep{jain2010}. Given the iterative nature of the algorithm, the initial partitioning is often important because the algorithm might arrive at a suboptimal clustering where the $WCSS$ cannot be further reduced by moving any feature to another cluster, despite a better solution existing \citep[i.e., a local minimum;][]{timmerman2013}. It is, therefore often recommended to run the k-means clustering with several different starting positions.

In our case we entered the participants' PC-scores from the feature reduction step into the k-means algorithm. Because we did know the underlying number of clusters within our sample, we calculated the cluster solutions for $k =$ 2 through 10. To avoid local minima we used 100 random initial centroid positions for each run. Each of the 9 cluster solutions converged within the iteration limit. In the next step we will then evaluate which of the extracted cluster solutions offers the best fit with the data.
